---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->
#joinXL 
This R package performs 'SQL' type joins and minus operations on 'Excel' files.
'SQL' is a programming language, which manipulates relational tables within a database.
joinXL provides 5 functions, which perform relational joins and a minus operation 
on 'Excel' files living on your hard drive.    
**innerjoinXL(), leftjoinXL(), rightjoinXL(), fulljoinXL(), and minusXL()**

#Installation
joinXL is only on github as of August 22, 2016, but  has been submitted to CRAN.
```{r eval=FALSE}
devtools::install_github("yvonneglanville/joinXL")
```


#Usage
The 'Excel' files must have a related **key** (column) for a join or minus operation to be performed.
All join operations require one file to be designated as the base file because the base file row values
take precedence over those of the join file.

Once the joinXL package is installed (described below), **each join can be called at the console by typing the function 
name with parenthesis after the name and hitting enter. There is no need to put file information within the parenthesis because the function will open a dialog box.** Choose 2 files, and then follow the console prompts to designate a base file, and choose the common column in the base and join files. A column is chosen from each file to accomadate inconsistent naming conventions. A joined file is then output to the working directory.



##minusXL() 
The minusXL() function allows the direct comparison of the rows of two tables. This function is called in the same manner as the joins. A dialog window for the input of two files opens, and then there are prompts at the console for the designation of a source file and and the common column in each file. Two operations are then performed on the tables.

**Operation 1.** source file-minus-target file outputs rows found in source file but missing from target    
**Operation 2.** target file-minus-source file outputs rows found in target file but missing from base

##If the source and the target '.xlsx' files are identical, then the output files will be empty.

**Output at the console**   
1. Preview of source file-minus-target file   
2. Preview of target file-minus-source file   
3. Preview of source file-minus-target file + target file-minus-source file   

**Output in working directory**   
'sourcemMINUStarget.xlsx'   
'targetMINUSsource.xlsx'   
'rowsNOTduplicated.xlsx'   






```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
```
